---
title: "PUD (13 vars): Correct the k-anon violations (2)"
#author: "Nimesh Patel"
latex_engine: xelatex
geometry: top = 1cm, bottom=1cm, left=2cm, right=1cm
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
  tables:
    style: Table
    layout: autofit
    width: 4.5
    caption:
      style: Table Caption
      pre: 'Table '
      sep: ': '
    conditional:
      first_row: true
      first_column: false
      last_row: false
      last_column: false
      no_hband: false
      no_vband: true
classoption: landscape
---
\pagenumbering{gobble}
```{r setup, include=FALSE}

cat(toString(Sys.time()))
source("../R/functions.R")
library(flextable)
library(arrow)
library(sdcMicro)

report_dir = "../reports"
out_dir = "../output"
data_dir = "../data/raw"

KANON_LEVEL <- 5
KANON_LEVEL
LDIV_LEVEL <- 2
LDIV_LEVEL

quasi_identifiers = c("sex","age_group","race","ethnicity")
confidential_attributes = c("pos_spec_dt")

file_name <- "COVID_Cases_Public_Limited_20210422_13_vars.parquet"

suppressed_file_name = paste(out_dir,"/",file_name,".suppressed.csv",sep="")
public_file_name = paste(data_dir,"/",file_name,sep="")
print(public_file_name)
df = read_parquet(public_file_name, as_data_frame = TRUE)
data <- data.frame(df)
knitr::opts_chunk$set(echo = TRUE)
```
## Summarize existing utility
```{r ft.align="left"}
quick <- quick_summary(data, label="all_fields", qis=quasi_identifiers)

```
## Summarize existing utility
```{r ft.align="left"}
utility <- summarize_utility(data, quasi_identifiers)

```

## Recoding all the "NA" (already suppressed), Missings and Unknowns to NA for purposes of k-anonymity

```{r ft.align="left"}
data_na <- recode_to_na(data,quasi_identifiers,BLANK_CATEGORIES)
```
## Set up sdcMicro object, using data_na and change alpha=c(0)) to alpha=c(1)) to correct a k-anon violations (2) problem: 
  *Output from using data before recode to na and alpha=c(0)): 
  *a k-anon violations ( 2 ) for k=( 5 ) and quasi-identifiers ( race ethnicity sex age_group ). 
  *If greater than zero violations, then here's 5 violations. race ethnicity sex age_group fk
  *15600394 NA NA NA NA 2
  *15600393 NA NA NA NA 2

```{r ft.align="left"}
sdcObj <- createSdcObj(dat=data_na,
                       keyVars=quasi_identifiers,
                       numVars=NULL,
                       weightVar=NULL,
                       hhId=NULL,
                       strataVar=NULL,
                       pramVars=NULL,
                       excludeVars=NULL,
                       seed=0,
                       randomizeRecords=FALSE,
                       alpha=c(1))

# print to confirm observations, num variales, quasis, quasi describes, and risk info
sdc_print(sdcObj, KANON_LEVEL)
```
## Print out the number of violations and a sample: k-anon violations should be zero
```{r ft.align="left"}
fk = summarize_violations(data_na, sdcObj, KANON_LEVEL, quasi_identifiers)

```



