---
latex_engine: xelatex
geometry: top = 1cm, bottom=1cm, left=1.5cm, right=2cm
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
  tables:
    style: Table
    layout: autofit
    width: 1.0
    caption:
      style: Table Caption
      pre: 'Table '
      sep: ': '
    conditional:
      first_row: true
      first_column: false
      last_row: false
      last_column: false
      no_hband: false
      no_vband: true
---
\pagenumbering{gobble}
```{r setup, include=FALSE}
source("../R/functions.R")

library(arrow)
library(flextable)
#data_dir = "../data/raw"
data_dir = "../data/raw"

file_name <- list.files(data_dir, "COVID_Cases_Public_Limited")

detailed_file_name = paste(data_dir,"/",file_name,sep="")

#change me to whatever version you want to display
date_from_filename <- gsub(".parquet", "", gsub("COVID_Cases_Public_Limited_","", file_name))
    
    
#change me to whatever version you want to display
dataset_version <- paste0(substr(date_from_filename, 1, 4),'-',
                          substr(date_from_filename,5,6), '-', 
                          substr(date_from_filename,7,8))
dataset_version <- "2021-10-12"

quasi_identifiers = c("race_ethnicity_combined","sex","age_group")

data <- read_parquet(detailed_file_name, as_data_frame = TRUE)

quick <- quick_summary(data, label="all_fields", qis=quasi_identifiers, print=FALSE)

utility <- summarize_utility(data, quasi_identifiers, print=FALSE)

knitr::opts_chunk$set(echo = FALSE)
set_flextable_defaults(
  fonts_ignore=TRUE
)
```

# [COVID-19 Case Surveillance Public Use Data](https://data.cdc.gov/Case-Surveillance/COVID-19-Case-Surveillance-Public-Use-Data/vbim-akqf) Utility Summary

Users should consider the level of completeness, including suppression levels when planning their analyses and use of public datasets. Privacy protections will suppress field values to reduce reidentification risks. Completeness varies by jurisdiction (i.e., state, local, and territorial) and time period. Variables are consistently coded to the value "Unknown" when jurisdictions specify in the case data submitted to CDC that the value is unknown, the value "Missing" when jurisdictions do not provide a value, and the value "NA" when the value is suppressed as part of privacy protections.  

Dataset version: `r dataset_version`

Total records in dataset: `r commas(nrow(data))`

## Quick Summary
```{r ft.align="left"}
quick_df <- data.frame(cbind(" "=row.names(quick),quick))
colnames(quick_df) <- c(" ",colnames(quick))
ft1 <- theme_zebra(flextable(quick_df), odd_header="transparent", odd_body="#dce4f4",even_body="transparent")
ft1 <- align(ft1, j=c("all_fields","quasi_fields"), align="right")
ft1 <- line_spacing(ft1, space=1, part="all")
ft1 <- fontsize(ft1, size=8, part="all")
ft1 <- padding(ft1, padding.top=0, part="all")
ft1 <- hrule(ft1, rule="atleast", part="all")
ft1 <- height_all(ft1, 2, part="all")
ft1 <- hline(ft1, part="all")
ft1 <- vline(ft1, part="all")
ft1 <- border_outer(ft1, part="all")
#ft1 <- autofit(ft1, add_h=100)
ft1 <- autofit(ft1)
ft1
```

## Field-level Utility Summary
```{r ft.align="left"}
ft2 <- theme_zebra(flextable(cbind(" "=row.names(utility),utility)), odd_header="transparent", odd_body="#dce4f4",even_body="transparent")
ft2 <- autofit(ft2)
ft2 <- align(ft2, j=c("suppressed","suppressed_percent","missing","missing_percent"), align="right")
#ft2 <- height(ft2, height=1)
#ft2 <- hrule(ft2, rule="auto")
ft2 <- line_spacing(ft2, space=1, part="all")
ft2 <- fontsize(ft2, size=8, part="all")
ft2 <- hline(ft2, part="all")
ft2 <- vline(ft2, part="all")
ft2 <- border_outer(ft2, part="all")
ft2
```

